Windowed least-squares model reduction for dynamical systems

### 2019-10-08 main.tex
proofing

\address[a]{Sandia National Laboratories,  Livermore, CA}
\begin{abstract}
This work proposes a \methodNameLower\ (\methodAcronym) approach for model-reduction
	of dynamical systems. The proposed approach sequentially minimizes the
	time-continuous full-order-model residual within a low-dimensional space--time trial
	subspace over time windows. The approach comprises a generalization 
of existing model reduction approaches, as particular instances of
  the methodology recover the Galerkin,
	least-squares Petrov--Galerkin (LSPG), and space–time LSPG approaches. In
	addition, the approach
	addresses key deficiencies in existing model-reduction
	techniques, e.g., the dependence of the (space--time) LSPG approach on the
	time discretization and the exponential growth in time exhibited by \textit{a posteriori}
	error bounds for both the Galerkin and LSPG approaches.  We consider two types of
	space--time trial
	subspaces within the proposed approach: one that associates with spatial projection only, and one that 
  associates with space--time projection. For each type of trial
	subspace, we consider two different solution techniques: direct (i.e.,
	discretize-then-optimize) and indirect (i.e., optimize-then-discretize).
	Numerical experiments conducted using trial subspaces characterized by spatial
	projection only demonstrate that the \methodAcronym\
	approach can yield more accurate solutions with lower space--time residuals than
	the Galerkin and LSPG approaches.

@@

Projection-based reduced-order models (ROMs) comprise one such approximation
strategy. First, these techniques execute a computationally expensive
\textit{offline} stage that computes a low-dimensional \textit{trial subspace} on
which the dynamical-system state can be well approximated (e.g., by computing
state ``snapshots'' at different time and parameter instances, by solving
Lyapunov equations). Second, these methods execute an inexpensive
\textit{online} stage during which they compute approximations to the
dynamical-system trajectory that reside on this trial subspace via, e.g., 
residual minimization.

@@

The paper proceeds as follows: Section~\ref{sec:math}
outlines the mathematical setting for the full-order model, along with the Galerkin, LSPG, and
space--time LSPG approaches. Section~\ref{sec:tclspg} outlines the proposed \methodAcronym\
approach. Section~\ref{sec:numerical_techniques}
outlines numerical techniques for solving \methodAcronymROMs, including both direct and
indirect methods. Section~\ref{sec:analysis} provides
equivalence conditions and error analysis for \methodAcronymROMs.
Section~\ref{sec:numerical_experiments} presents numerical experiments.
Section~\ref{sec:conclude} provides conclusions and perspectives.
We denote vector-valued functions with italicized bold symbols (e.g., $\boldsymbol
x$), vectors with standard bold symbols (e.g., $\mathbf{x}$), 
matrices with capital bold symbols (e.g., $\mathbf{X} \equiv \begin{bmatrix}
\mathbf{x}_1 & \cdots & \mathbf{x}_r\end{bmatrix}$), and spaces with
calligraphic symbols (e.g., $\mathcal{X}$). We additionally denote differentiation of a time-dependent 
function with respect to with the $\dot\ $ operator.
%Model reduction of nonlinear dynamical systems is a growing research area.
%The least-squares Petrov-Galerkin (LSPG) approach has gained popularity as a model reduction technique for nonlinear dynamical systems. Derived at the fully discrete level, the LSPG approach solves a least-squares problem that minimizes the fully discrete residual of the reduced-order model at each time-step. LSPG has been shown to be an effective formulation for nonlinear model reduction on complex problems of interest.
%\begin{comment} 
\section{Mathematical formulation}\label{sec:math}
	We begin by providing the formulation for the full-order model,
	followed by a description of standard model-reduction methods
	classified according to the type of trial subspace they employ.
	\subsection{Full-order model}\label{sec;FOM}
We consider the full-order model to be a dynamical system expressed as a
	system of ordinary differential equations (ODEs)
\begin{equation}\label{eq:FOM}
 \stateFOMDotArg{t}  = \velocity(\stateFOMArg{}{t},t), \qquad \stateFOM(0) =
	\stateFOMIC,\qquad t \in [0,T]
\end{equation}
where  $\stateFOM: [0,T]
	\rightarrow  \RR{\fomdim}$ with $\stateFOM: \timeDummy \mapsto
	\stateFOM(\timeDummy) $ and $\stateFOMDot \equiv {d \stateFOM}/{d\tau}$
	denotes the state implicitly defined as the solution to initial value
	problem~\eqref{eq:FOM}, 
$T \in \mathbb{R}^+$ denotes the final time, 
 $\stateFOMIC \in \mathbb{R}^{\fomdim}$ denotes the initial condition,
	and $\velocity: \mathbb{R}^{\fomdim} \times [0,T] \rightarrow
	\mathbb{R}^{\fomdim}$ with $(\stateyDiscrete,\timeDummy) \mapsto
	\velocity(\stateyDiscrete,\timeDummy)$ denotes the velocity, which is possibly
	nonlinear in its first argument. For subsequent exposition, we introduce
	$\timeSpace$ to denote the set of real-valued functions acting on the time
	domain (i.e., $\timeSpace = \{f\,|\,f:[0,T]\rightarrow\RR{}\}$); the
	state can be expressed equivalently as $\stateFOM \in \RR{\fomdim} \otimes
	\timeSpace $.  We refer to the initial value problem defined in
	Eq.~\eqref{eq:FOM} as the ``full-order model" (FOM) ODE. We note that
	although the problem of interest described in the introduction corresponds
	to a parameterized dynamical system, we suppress dependence of the FOM ODE
	\eqref{eq:FOM} on such parameters for notational convenience, as this work
	focuses on devising a model-reduction approach applicable to a specific parameter instance.

@@

\subsection{Reduced-order models}
Projection-based ROMs generate approximate solutions to the FOM
	ODE~\eqref{eq:FOM} by approximating the state in a low-dimensional trial
	subspace. Two types of space--time trial subspaces are commonly used for
	this purpose: 
\begin{enumerate} 
	\item \textit{Subspaces that reduce only the spatial dimension of the full-order
		model (\spatialAcronym)}. These trial spaces are characterized by a \textit{spatial projection} process and are employed in classic model reduction approaches, e.g., Galerkin and LSPG. %These 
%	trial subspaces are commonly used
%		in model-reduction methods based on spatial projection followed by time
%		integration (e.g., the Galerkin and LSPG approaches).
	\item \textit{Subspaces that reduce both the spatial and temporal dimensions of the full-order
		model (\spaceTimeAcronym)}. 
 These trial subspaces are commonly employed in
		model-reduction methods based on space--time projection (e.g., space--time
		LSPG).
\end{enumerate}
 We now describe these two types of space--time trial subspaces and their
	application to the Galerkin, LSPG, and space--time LSPG approaches.

@@

\subsubsection{Galerkin approach}
The Galerkin approach reduces the number of equations in
Eq.~\eqref{eq:g_truncation} by enforcing orthogonality of the 
residual to the spatial trial subspace in the (semi-)inner product induced by the 
positive (semi-)definite $\fomdim\times\fomdim$ matrix $\stweightingMatArg{}\equiv \stweightingMatOneTArg{ }
\stweightingMatOneArg{ }$ (commonly set to
$\stweightingMatArg{}=\mathbf{I}$), i.e.,
\begin{equation}\label{eq:g_truncation2}
\genstateGalerkinDotArg{}{t} = \massArgnt{n}^{-1} \basisspace^T \stweightingMatArg{} \velocity(\basisspace
\genstateGalerkinArg{}{t} + \stateIntercept,t), \qquad \genstateGalerkin(0) = \genstateIC, \qquad t \in [0,T],
\end{equation}
where $\massArgnt{ } \equiv \basisspace^T \stweightingMatArg{ } \basisspace$
denotes the $K\times K$ positive definite mass matrix.
As demonstrated in Ref.~\cite{carlberg_lspg_v_galerkin}, the Galerkin approach
can be viewed alternatively as a residual-minimization
method, as the Galerkin ODE~\eqref{eq:g_truncation2} is equivalent
to
\begin{equation}\label{eq:GalOptimal}
\genstateGalerkinDotArg{}{t} = \underset{ \genstateyDiscrete \in \RR{\romdim}
	}{\text{arg\,min}} \| \basisspace \hat{\stateyDiscrete} -
	\velocity(\basisspace \genstateGalerkinArg{}{t} + \stateIntercept,t)
	\|_{\stweightingMatArg{}}^2, \qquad \genstate(0) = \genstateIC,
	\qquad t \in [0,T],
%\hat{\mathbf{v}}, %\qquad \genstateGalerkin(0) = \basisspace^T(\stateFOMIC -\stateIntercept), \qquad t \in [0,T].
\end{equation}
where $\|\mathbf{x}\|_{\stweightingMatArg{}}\equiv
\sqrt{\mathbf{x}^T\stweightingMatArg{}\mathbf{x}}$.
Thus, the computed velocity $\genstateGalerkinDotArg{}{t}$ minimizes the
FOM ODE residual evaluated at the state $\basisspace
\genstateGalerkin(t) + \stateIntercept$ and time instance $t$ over the spatial trial
subspace $\trialspace$.
%To perform dimension reduction, 
%To perform dimension reduction, the governing equations are projected by a ``test" basis,
%\begin{equation}\label{eq:g_truncation}
%\testbasis^T \basisspace \frac{d \genstate}{dt} = \testbasis^T \velocity(\basisspace \genstate + \state_0,t; \param ), \qquad t \in [0,T],
%\end{equation}
%The most common choice is to set the test space and trial space to be equivalent, which results in the Galerkin reduced-order model,
%\begin{equation}\label{eq:g_truncation}
%\frac{d \genstate}{dt} = \basisspace^T \velocity(\basisspace \genstate + \state_0,t; \param ), \qquad t \in [0,T],
%\end{equation}
\subsubsection{The Least-squares Petrov--Galerkin approach}
Despite its time-instantaneous residual-minimization optimality property 
\eqref{eq:GalOptimal}, the Galerkin approach can yield inaccurate solutions,
particularly if the velocity is not self-adjoint or
is nonlinear. 
Least-squares Petrov--Galerkin (LSPG)
~\cite{carlberg_lspg_v_galerkin,carlberg_thesis,carlberg_gnat,bui_unsteady,bui_thesis}
was developed
as an alternative method that exhibits several advantages over the Galerkin approach.
Rather than minimize the (time-continuous) FOM ODE residual at a time instance (as in Galerkin), LSPG minimizes
the (time-discrete) FOM O$\Delta$E residual 
(i.e., the residual arising after applying time discretization to the FOM ODE)
over a time step.
We now describe the LSPG approach in the case of linear multistep methods;
Ref.~\cite{carlberg_lspg_v_galerkin} also presents LSPG for Runge--Kutta
schemes.

Without loss of generality, we introduce a uniform time
grid characterized by time step $\Delta t$ and time instances
$t^n = n\Delta
t$, $n=0,\ldots,N_t$.
Applying a linear multistep method to discretize the FOM ODE \eqref{eq:FOM}
with this time grid
yields the FOM O$\Delta$E, which computes the sequence of discrete
solutions
$\stateFOMDiscreteArg{n}( \approx \stateFOM(t^n))$, $n=1,\ldots,N_t$
as the implicit solution to the system of algebraic equations
\begin{equation}\label{eq:lms}
\residLMSArg{n}
	(\stateFOMDiscreteArg{n};\stateFOMDiscreteArg{n-1},\ldots,\stateFOMDiscreteArg{n-k^n})
	= \bz,\qquad n=1,\ldots,N_t,
\end{equation}
where  $k^n$ denotes the number of steps employed by the scheme at the $n$th
time instance and 
$\residLMSArg{n}$ denotes the FOM O$\Delta$E residual defined as
\begin{align*}
\residLMSArg{n} &: (\stateyDiscreteArgnt{n};\stateyDiscreteArgnt{n-1},\ldots,\stateyDiscreteArgnt{n-k^n}) \mapsto  \frac{1}{\Delta t} \sum_{j=0}^{k^n} \alpha^n_j \stateyDiscreteArgnt{n-j} -  \sum_{j=0}^{k^n} \beta^n_j \velocity(\stateyDiscreteArgnt{n-j},t^{n-j}),
\\
&: \RR{\fomdim} \otimes \RR{k^n + 1} \rightarrow \RR{\fomdim},
\end{align*} 
Here, $\alpha^n_j,\beta^n_j\in\RR{}$, $j=0,\ldots,k^n$ are coefficients
that define the linear multistep method at the $n$th time instance.
%\begin{equation}\label{eq:coarse_implicit_euler_0}
%\residLMS \frac{\stateFOMDiscreteArg{n} - \stateFOMDiscreteArg{n-1}  }{\Delta t} - \velocity(\stateFOMDiscreteArg{n}) = \mathbf{0}, \qquad n = 1,2,\ldots,N_t,
%\end{equation}
%the algebraic FOM O$\Delta$E obtained after an implicit Euler temporal discretization is given by,
%\begin{equation}\label{eq:coarse_implicit_euler_0}
%\frac{\stateFOMDiscreteArg{n} - \stateFOMDiscreteArg{n-1}  }{\Delta t} - \velocity(\stateFOMDiscreteArg{n}) = \mathbf{0}, \qquad n = 1,2,\ldots,N_t,
%\end{equation}
%where $\stateFOMDiscreteArg{n} \approx \stateFOM(t^n)$ denotes the discrete approximation to the state at the $n${th} time instance.

@@

\spaceTimeAcronym\ trial subspaces approximate the FOM ODE solution
trajectory
	$\stateFOM\in\RR{N}\otimes \timeSpace$ with an approximation that resides in an
	affine space--time trial subspace of dimension $\stdim\ll\fomdim$, i.e., 
	$\approxstate\in \stspaceST$ with $\dim(\stspaceST) =\stdim $, where
\begin{equation}\label{eq:sttrialspace_def}
 \stspaceST \defeq 
	\Range{\stbasis} + 
	\stateFOMIC\otimes\onesFunction
	\subseteq \RR{N} \otimes \timeSpace
\end{equation}
and $\stbasis\in\RR{N \times \stdim}\otimes \timeSpace$ with 
$\stbasis:\timeDummy\mapsto\stbasis(\timeDummy)$ and $\stbasis(0) =
\boldsymbol 0$
denotes the space--time trial basis.\footnote{For simplicity, we consider the case where the trial space is homogeneous on the boundary.} 
Thus, at any time instance $t\in[0,T]$, ROMs that employ the
\spaceTimeAcronym\ trial subspace approximate the state as
\begin{equation}\label{eq:stapprox1}
 \stateFOMArg{}{t} \approx \approxstateArg{}{t}  = \stbasisArg{}{t} \stgenstate + \stateFOMIC,
\end{equation}
where $\stgenstateArg{} \in \RR{\stdim}$ denote the space--time generalized coordinates. 
Critically, comparing the approximations arising from \spatialAcronym\ and
\spaceTimeAcronym\ trial subspaces in Eqs.~\eqref{eq:affine_trialspace} and \eqref{eq:stapprox1}, respectively,
highlights that the former approximation associates with time-dependent
generalized coordinates, while the latter approximation associates with a
time-dependent basis matrix.

>>>>

\address[a]{Sandia National Laboratories,  Livermore, CA}
\begin{abstract}
This work proposes a \methodNameLower\ (\methodAcronym) approach for model-reduction
	of dynamical systems. The proposed approach sequentially minimizes the
	time-continuous full-order-model residual within a low-dimensional space--time trial
	subspace over time windows. The approach comprises a generalization 
of existing model reduction approaches, as particular instances of
  the methodology recover the Galerkin,
	least-squares Petrov--Galerkin (LSPG), and space–time LSPG approaches. In
	addition, the approach
	addresses key deficiencies in existing model-reduction
	techniques, e.g., the dependence of the (space--time) LSPG approach on the
	time discretization and the exponential growth in time exhibited by \textit{a posteriori}
	error bounds for both the Galerkin and LSPG approaches.  We consider two types of
	space--time trial
	subspaces within the proposed approach: one that reduces only the spatial dimension of the full-order model, and one that reduces both the spatial and temporal dimensions of the full-order model. For each type of trial
	subspace, we consider two different solution techniques: direct (i.e.,
	discretize-then-optimize) and indirect (i.e., optimize-then-discretize).
	Numerical experiments conducted using trial subspaces characterized by spatial
	dimension reduction demonstrate that the \methodAcronym\
	approach can yield more accurate solutions with lower space--time residuals than
	the Galerkin and LSPG approaches.

@@

Projection-based reduced-order models (ROMs) comprise one such approximation
strategy. First, these techniques execute a computationally expensive
\textit{offline} stage that computes a low-dimensional \textit{trial subspace} on
which the dynamical-system state can be well approximated (e.g., by computing
state ``snapshots'' at different time and parameter instances, by solving
Lyapunov equations). Second, these methods execute an inexpensive
\textit{online} stage during which they compute approximations to the
dynamical-system trajectory that reside on this trial subspace via, e.g., 
projection of the full-order model or residual minimization.

@@

The paper proceeds as follows: Section~\ref{sec:math}
outlines the mathematical setting for the full-order model, along with the Galerkin, LSPG, and
space--time LSPG approaches. Section~\ref{sec:tclspg} outlines the proposed \methodAcronym\
approach. Section~\ref{sec:numerical_techniques}
outlines numerical techniques for solving \methodAcronymROMs, including both direct and
indirect methods. Section~\ref{sec:analysis} provides
equivalence conditions and error analysis for \methodAcronymROMs.
Section~\ref{sec:numerical_experiments} presents numerical experiments.
Section~\ref{sec:conclude} provides conclusions and perspectives.
We denote vector-valued functions with italicized bold symbols (e.g., $\boldsymbol
x$), vectors with standard bold symbols (e.g., $\mathbf{x}$), 
matrices with capital bold symbols (e.g., $\mathbf{X} \equiv \begin{bmatrix}
\mathbf{x}_1 & \cdots & \mathbf{x}_r\end{bmatrix}$), and spaces with
calligraphic symbols (e.g., $\mathcal{X}$). We additionally denote differentiation of a time-dependent 
function with respect to time with the $\dot\ $ operator.
%Model reduction of nonlinear dynamical systems is a growing research area.
%The least-squares Petrov-Galerkin (LSPG) approach has gained popularity as a model reduction technique for nonlinear dynamical systems. Derived at the fully discrete level, the LSPG approach solves a least-squares problem that minimizes the fully discrete residual of the reduced-order model at each time-step. LSPG has been shown to be an effective formulation for nonlinear model reduction on complex problems of interest.
%\begin{comment} 
\section{Mathematical formulation}\label{sec:math}
	We begin by providing the formulation for the full-order model,
	followed by a description of standard model-reduction methods
	classified according to the type of trial subspace they employ.
	\subsection{Full-order model}\label{sec;FOM}
We consider the full-order model to be a dynamical system expressed as a
	system of ordinary differential equations (ODEs)
\begin{equation}\label{eq:FOM}
 \stateFOMDotArg{t}  = \velocity(\stateFOMArg{}{t},t), \qquad \stateFOM(0) =
	\stateFOMIC,\qquad t \in [0,T],
\end{equation}
where  $\stateFOM: [0,T]
	\rightarrow  \RR{\fomdim}$ with $\stateFOM: \timeDummy \mapsto
	\stateFOM(\timeDummy) $ and $\stateFOMDot \equiv {d \stateFOM}/{d\tau}$
	denotes the state implicitly defined as the solution to initial value
	problem~\eqref{eq:FOM}, 
$T \in \RRplus$ denotes the final time, 
 $\stateFOMIC \in \mathbb{R}^{\fomdim}$ denotes the initial condition,
	and $\velocity: \mathbb{R}^{\fomdim} \times [0,T] \rightarrow
	\mathbb{R}^{\fomdim}$ with $(\stateyDiscrete,\timeDummy) \mapsto
	\velocity(\stateyDiscrete,\timeDummy)$ denotes the velocity, which is possibly
	nonlinear in its first argument. For subsequent exposition, we introduce
	$\timeSpace$ to denote the set of (sufficiently smooth) real-valued functions acting on the time
	domain (i.e., $\timeSpace = \{f\,|\,f:[0,T]\rightarrow\RR{}\}$); the
	state can be expressed equivalently as $\stateFOM \in \RR{\fomdim} \otimes
	\timeSpace $.  We refer to the initial value problem defined in
	Eq.~\eqref{eq:FOM} as the ``full-order model" (FOM) ODE. We note that
	although the problem of interest described in the introduction corresponds
	to a parameterized dynamical system, we suppress dependence of the FOM ODE
	\eqref{eq:FOM} on such parameters for notational convenience, as this work
	focuses on devising a model-reduction approach applicable to a specific parameter instance.

@@

\subsection{Reduced-order models}
Projection-based ROMs generate approximate solutions to the FOM
	ODE~\eqref{eq:FOM} by approximating the state in a low-dimensional trial
	subspace. Two types of space--time trial subspaces are commonly used for
	this purpose:\footnote{For both spatial and space--time ROMs of dynamical systems, all trial subspaces are, strictly speaking, space--time subspaces.} 
\begin{enumerate} 
	\item \textit{Subspaces that reduce only the spatial dimension of the full-order
		model (\spatialAcronym)}. These trial subspaces are characterized by a \textit{spatial projection} and are employed in classic model reduction approaches, e.g., Galerkin and LSPG. %These 
%	trial subspaces are commonly used
%		in model-reduction methods based on spatial projection followed by time
%		integration (e.g., the Galerkin and LSPG approaches).
	\item \textit{Subspaces that reduce both the spatial and temporal dimensions of the full-order
		model (\spaceTimeAcronym)}.
These trial subspaces are characterized by a \textit{space--time} projection and are employed in space--time 
model reduction approaches (e.g., space--time Galerkin~\cite{benner_st}, space--time LSPG~\cite{choi_stlspg}). 
% These trial subspaces are commonly employed in
%		model-reduction methods based on space--time projection (e.g., space--time
%		LSPG).
\end{enumerate}
 We now describe these two types of space--time trial subspaces and their
	application to the Galerkin, LSPG, and space--time LSPG approaches.

@@

\subsubsection{Galerkin approach}
The Galerkin approach reduces the number of equations in
Eq.~\eqref{eq:g_truncation} by enforcing orthogonality of the 
residual to the spatial trial subspace in the (semi-)inner product induced by the 
positive (semi-)definite $\fomdim\times\fomdim$ matrix $\stweightingMatArg{}\equiv \stweightingMatOneTArg{ }
\stweightingMatOneArg{ }$ (commonly set to
$\stweightingMatArg{}=\mathbf{I}$), i.e.,
\begin{equation}\label{eq:g_truncation2}
\genstateGalerkinDotArg{}{t} = \massArgnt{n}^{-1} \basisspace^T \stweightingMatArg{} \velocity(\basisspace
\genstateGalerkinArg{}{t} + \stateIntercept,t), \qquad \genstateGalerkin(0) = \genstateIC, \qquad t \in [0,T],
\end{equation}
where $\massArgnt{ } \equiv \basisspace^T \stweightingMatArg{ } \basisspace$
denotes the $K\times K$ positive definite mass matrix.
As demonstrated in Ref.~\cite{carlberg_lspg_v_galerkin}, the Galerkin approach
can be viewed alternatively as a residual-minimization
method, as the Galerkin ODE~\eqref{eq:g_truncation2} is equivalent
to
\begin{equation}\label{eq:GalOptimal}
\genstateGalerkinDotArg{}{t} = \underset{ \genstateyDiscrete \in \RR{\romdim}
	}{\text{arg\,min}} \| \basisspace \hat{\stateyDiscrete} -
	\velocity(\basisspace \genstateGalerkinArg{}{t} + \stateIntercept,t)
	\|_{\stweightingMatArg{}}^2, \qquad \genstate(0) = \genstateIC,
	\qquad t \in [0,T],
%\hat{\mathbf{v}}, %\qquad \genstateGalerkin(0) = \basisspace^T(\stateFOMIC -\stateIntercept), \qquad t \in [0,T].
\end{equation}
where $\|\mathbf{x}\|_{\stweightingMatArg{}}\equiv
\sqrt{\mathbf{x}^T\stweightingMatArg{}\mathbf{x}}$.
Thus, the computed velocity $\genstateGalerkinDotArg{}{t}$ minimizes the
FOM ODE residual evaluated at the state $\basisspace
\genstateGalerkin(t) + \stateIntercept$ and time instance $t$ over the spatial trial
subspace $\trialspace$.
%To perform dimension reduction, 
%To perform dimension reduction, the governing equations are projected by a ``test" basis,
%\begin{equation}\label{eq:g_truncation}
%\testbasis^T \basisspace \frac{d \genstate}{dt} = \testbasis^T \velocity(\basisspace \genstate + \state_0,t; \param ), \qquad t \in [0,T],
%\end{equation}
%The most common choice is to set the test space and trial space to be equivalent, which results in the Galerkin reduced-order model,
%\begin{equation}\label{eq:g_truncation}
%\frac{d \genstate}{dt} = \basisspace^T \velocity(\basisspace \genstate + \state_0,t; \param ), \qquad t \in [0,T],
%\end{equation}
\subsubsection{The least-squares Petrov--Galerkin approach}
Despite its time-instantaneous residual-minimization optimality property 
\eqref{eq:GalOptimal}, the Galerkin approach can yield inaccurate solutions,
particularly if the velocity is not self-adjoint or
is nonlinear. 
Least-squares Petrov--Galerkin (LSPG)
~\cite{carlberg_lspg_v_galerkin,carlberg_thesis,carlberg_gnat,bui_unsteady,bui_thesis}
was developed
as an alternative method that exhibits several advantages over the Galerkin approach.
Rather than minimize the (time-continuous) FOM ODE residual at a time instance (as in Galerkin), LSPG minimizes
the (time-discrete) FOM O$\Delta$E residual 
(i.e., the residual arising after applying time discretization to the FOM ODE)
over a time step.
We now describe the LSPG approach in the case of linear multistep methods;
Ref.~\cite{carlberg_lspg_v_galerkin} also presents LSPG for Runge--Kutta
schemes.

Without loss of generality, we introduce a uniform time
grid characterized by time step $\Delta t$ and time instances
$t^n = n\Delta
t$, $n=0,\ldots,N_t$.
Applying a linear multistep method to discretize the FOM ODE \eqref{eq:FOM}
with this time grid
yields the FOM O$\Delta$E, which computes the sequence of discrete
solutions
$\stateFOMDiscreteArg{n}( \approx \stateFOM(t^n))$, $n=1,\ldots,N_t$
as the implicit solution to the system of algebraic equations
\begin{equation}\label{eq:lms}
\residLMSArg{n}
	(\stateFOMDiscreteArg{n};\stateFOMDiscreteArg{n-1},\ldots,\stateFOMDiscreteArg{n-k^n})
	= \bz,\qquad n=1,\ldots,N_t,
\end{equation}
where  $k^n$ denotes the number of steps employed by the scheme at the $n$th
time instance and 
$\residLMSArg{n}$ denotes the FOM O$\Delta$E residual defined as
\begin{align*}
\residLMSArg{n} &: (\stateyDiscreteArgnt{n};\stateyDiscreteArgnt{n-1},\ldots,\stateyDiscreteArgnt{n-k^n}) \mapsto  \frac{1}{\Delta t} \sum_{j=0}^{k^n} \alpha^n_j \stateyDiscreteArgnt{n-j} -  \sum_{j=0}^{k^n} \beta^n_j \velocity(\stateyDiscreteArgnt{n-j},t^{n-j}),
\\
&: \RR{\fomdim} \otimes \RR{k^n + 1} \rightarrow \RR{\fomdim}.
\end{align*} 
Here, $\alpha^n_j,\beta^n_j\in\RR{}$, $j=0,\ldots,k^n$ are coefficients
that define the linear multistep method at the $n$th time instance.
%\begin{equation}\label{eq:coarse_implicit_euler_0}
%\residLMS \frac{\stateFOMDiscreteArg{n} - \stateFOMDiscreteArg{n-1}  }{\Delta t} - \velocity(\stateFOMDiscreteArg{n}) = \mathbf{0}, \qquad n = 1,2,\ldots,N_t,
%\end{equation}
%the algebraic FOM O$\Delta$E obtained after an implicit Euler temporal discretization is given by,
%\begin{equation}\label{eq:coarse_implicit_euler_0}
%\frac{\stateFOMDiscreteArg{n} - \stateFOMDiscreteArg{n-1}  }{\Delta t} - \velocity(\stateFOMDiscreteArg{n}) = \mathbf{0}, \qquad n = 1,2,\ldots,N_t,
%\end{equation}
%where $\stateFOMDiscreteArg{n} \approx \stateFOM(t^n)$ denotes the discrete approximation to the state at the $n${th} time instance.

@@

\spaceTimeAcronym\ trial subspaces approximate the FOM ODE solution
trajectory
	$\stateFOM\in\RR{N}\otimes \timeSpace$ with an approximation that resides in an
	affine space--time trial subspace of dimension $\stdim\ll\fomdim$, i.e., 
	$\approxstate\in \stspaceST$ with $\dim(\stspaceST) =\stdim $, where
\begin{equation}\label{eq:sttrialspace_def}
 \stspaceST \defeq 
	\Range{\stbasis} + 
	\stateFOMIC\otimes\onesFunction
	\subseteq \RR{N} \otimes \timeSpace
\end{equation}
%and $\stbasis\in\RR{N \times \stdim}\otimes \timeSpace$ with 
and $\stbasis:\timeDummy\mapsto\stbasis(\timeDummy)$ with $\stbasis(0) =
\boldsymbol 0$
denotes the space--time trial basis. 
Thus, at any time instance $t\in[0,T]$, ROMs that employ the
\spaceTimeAcronym\ trial subspace approximate the state as
\begin{equation}\label{eq:stapprox1}
 \stateFOMArg{}{t} \approx \approxstateArg{}{t}  = \stbasisArg{}{t} \stgenstate + \stateFOMIC,
\end{equation}
where $\stgenstateArg{} \in \RR{\stdim}$ denotes the space--time generalized coordinates. 
Critically, comparing the approximations arising from \spatialAcronym\ and
\spaceTimeAcronym\ trial subspaces in Eqs.~\eqref{eq:affine_trialspace} and \eqref{eq:stapprox1}, respectively,
highlights that the former approximation associates with time-dependent
generalized coordinates, while the latter approximation associates with a
time-dependent basis matrix.
